---
title: "TrIdent"
author: "Jessie Maier, Jorden Rabasco"
output: 
  BiocStyle::html_document:
    toc: true
    toc_depth: 3
description: >
  TrIdent (Transduction Identification) is a reference-independent tool for the 
  automated detection, classification and characterization of read coverage 
  patterns in transductomics datasets. 
vignette: >
  %\VignetteIndexEntry{TrIdent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  out.width = "100%",
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, warning=FALSE}
library(TrIdent)
library(knitr)
library(patchwork)
library(ggplot2)
library(kableExtra)
```

# Introduction

**TrIdent**- **Tr**ansduction **Ident**ification

TrIdent consists of three main functions which should be run in the following 
order:

1. **`TrIdentClassifier()`**: Classifies contigs as 'Prophage-like', 'Sloping', 
'HighCovNoPattern', and 'NoPattern' using read coverage pattern-matching.
2. **`plotTrIdentResults()`**: Plots pattern-matching results of 
`TrIdentClassifier()`.
3. **`specializedTransductionID()`**: Searches contigs classified as 
Prophage-like by `TrIdentClassifier()` for associated specialized transduction 
events.

TrIdent automates the analysis of transductomics data by detecting, 
classifying, and characterizing read coverage patterns associated with 
potential transduction events. Transductomics, developed by Kleiner et al. 
(2020), is a DNA sequencing-based method for the detection and characterization 
of transduction events in pure cultures and complex communities. Transductomics 
relies on mapping sequencing reads from a viral-like particle (VLP)-fraction of 
a sample to contigs assembled from the metagenome (whole-community) of the same 
sample. Reads from bacterial DNA carried by VLPs will map back to the bacterial 
contigs of origin creating read coverage patterns indicative of ongoing 
transduction. **The read coverage patterns detected represent DNA being**
**actively carried or transduced by VLPs. The read coverage patterns do not**
**represent complete transduction events (i.e integration of transduced DNA**
**into new bacterial chromosomes).** 

**Reference:** Kleiner, M., Bushnell, B., Sanderson, K.E. et al. 
Transductomics: sequencing-based detection and analysis of transduced DNA in 
pure cultures and microbial communities. Microbiome 8, 158 (2020). 
<https://doi.org/10.1186/s40168-020-00935-5>

## Installation

## Bioconductor install
```{r, eval=FALSE}
if (!require("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}

BiocManager::install("TrIdent")
library(TrIdent)
```

## GitHub install
```{r, eval=FALSE}
BiocManager::install("jlmaier12/TrIdent")
library(TrIdent)
```

# Input Data

## Transductomics data
Transductomics allows for the identification of bacterial DNA being actively 
carried or transduced by VLPs. A transductomics dataset consists of two parts- 
metagenomes from the whole-community and VLP fractions of a sample. The 
whole-community fraction is generated by extracting and sequencing DNA from the 
whole sample. The VLP-fraction is generated by extraction and sequencing DNA of 
the ultra-purified VLPs in the sample. VLP ultra-purification is generally done 
using CsCl density-gradient ultracentrifugation. Additionally, it is very 
important that the VLP-fraction is treated with DNase to remove free DNA! After 
sequencing, reads from the whole-community fraction are assembled and both the 
whole-community and VLP-fraction reads are mapped to the assembly. **Read** 
**mapping should be performed using a high minimum identity (0.97 or higher)** 
**and random mapping of ambiguous reads.** The pileup files needed for TrIdent 
are generated using the .bam files produced during read mapping. 

**Deep sequencing of the whole-community and VLP-fractions is needed for** 
**transductomics! Sample preparation, sequencing procedures, and** 
**bioinformatics methods are detailed in** 
**[Kleiner et al.(2020)](https://microbiomejournal.biomedcentral.com/articles/10.1186/s40168-020-00935-5).**

## Pileup files
TrIdent detects read coverage patterns using a pattern-matching algorithm that 
operates on pileup files. A pileup file is a file format where each row 
summarizes the 'pileup' of reads at specific genomic locations. Pileup files 
can be used to generate a rolling mean of read coverages and associated base 
pair positions across a metagenome assembly which reduces data size while 
preserving read coverage patterns. **TrIdent requires that input pileups files 
be generated using a 100 bp window/bin size.**  

Some read mappers, like 
[BBMap](https://jgi.doe.gov/data-and-tools/software-tools/bbtools/bb-tools-user-guide/bbmap-guide/), 
will allow for the generation of pileup files in the 
[`bbmap.sh`](https://github.com/BioInfoTools/BBMap/blob/master/sh/bbmap.sh) 
command with the use of the `bincov` output with the `covbinsize=100` 
parameter/argument. **Otherwise, BBMap's** 
**[`pileup.sh`](https://github.com/BioInfoTools/BBMap/blob/master/sh/pileup.sh)** 
**can convert .bam files produced by any read mapper to pileup files** 
**compatible with TrIdent using the `bincov` output with `binsize=100`.**

TrIdent requires two pileup files from a transductomics dataset as input:

* A VLP-fraction pileup: Sequencing reads from a sample's ultra-purified 
VLP-fraction mapped to the sample's whole-community metagenome assembly.
* A whole-community pileup: Sequencing reads from a sample's whole-community 
mapped to the sample's whole-community metagenome assembly. 

**Remember- The data used for each pileup file must originate from the same** 
**sample. Pileup files must use a 100 bp window/bin size for the rolling** 
**mean.** 

**The input pileup files must have the following format:**

Dataframe with four columns:

* V1: Contig accession
* V2: Mapped read coverage values averaged over 100 bp windows
* V3: Starting position (bp) of each 100 bp window. Restarts from 0 at the 
start of each new contig.
* V4: Starting position (bp) of each 100 bp window. Does NOT restart at the 
start of each new contig.

```{r echo=FALSE}
data("WholeCommunitySamplePileup")
data("VLPFractionSamplePileup")
kable(head(VLPFractionSamplePileup), row.names = FALSE) %>%
  kable_styling(latex_options = "HOLD_position")
```

# TrIdentClassifier()

`TrIdentClassifier()` is the main function in TrIdent. This function filters 
contigs based on length and read coverage, performs pattern-matching to 
classify contigs, identifies highly active/abundant and heterogenously 
integrated Prophage-like elements, determines which contigs have high 
VLP-fraction:whole-community read coverage ratios, identifies start and stop 
positions and sizes of pattern-matches, and calculates slopes for Sloping 
pattern-matches.

Note that `TrIdentClassifier()` results should not be taken at face value and 
some manual curation is needed! We recommend careful examination of the output 
plots to determine if you agree with the `TrIdentClassifier()` classifications. 

## Function components

### Contig filtering
Contigs that are too short or have little to no read coverage are filtered out 
prior to pattern-matching. `TrIdentClassifier()` filters out contigs that do 
not have at least 10x coverage on a total of 5,000 bp across the whole contig. 
The read coverage filtering was done in this way to avoid filtering out long 
contigs with short Prophage-like patterns that might get removed if filtering 
was done with read coverage averages or medians. Additionally, contigs less 
than 30,000 bp are filtered out by default, however this can be changed with 
the `minContigLength` parameter. While contigs shorter than 30,000 bp may be 
poor quality and are often not long enough to capture complete transduction 
patterns, we allow users to set a minimum `minContigLength` of 25,000 bp. 
**If you would like to reduce the size of your input pileup files for** 
**TrIdent, consider pre-filtering your assembly for contigs greater than** 
**25,000 bp prior to read mapping!** 

**It is expected that the majority of contigs in the dataset are filtered out**
**due to low read coverage during this step! If very few contigs are filtered**
**out due to low read coverage, this may be an indication of DNA** 
**contamination in your VLP-fraction.** 

### Changing pileup windowSize
The input pileup files have 100 bp windows in which the mapped read coverage is 
averaged over. This small window size is needed to detect read coverage 
patterns associated with specialized transduction which may only span a few 
thousand basepairs. However, read coverage patterns associated with other types 
of transduction generally don't require the resolution that 100 bp windows 
provide. While users can use the 100 bp `windowSize` for `TrIdentClassifier()`, 
the processing time will be increased significantly and noisy data may 
interfere with pattern-matching. We find that the default 1,000 bp `windowSize` 
provides a nice balance between processing time and read coverage pattern 
resolution.  

### Pattern-matching
`TrIdentClassifier()` detects read coverage patterns in the VLP-fraction using 
a pattern-matching approach. Several predefined patterns, described below, are 
built using the specific length and read coverage values of the contig being 
assessed. Patterns are translated across each contig in 1,000 bp sliding 
windows and at each translation, a pattern-match score is calculated by taking 
the mean absolute difference of the VLP-fraction read coverage and the pattern 
values. The smaller the match-score, the better the pattern-match. After a 
pattern is fully translated across a contig, certain aspects of the pattern are 
changed (i.e. height, width, slope) and translation is repeated. This process 
of translation and pattern re-scaling is repeated until a large number of 
pattern variations are tested. After pattern-matching is complete, the pattern 
associated with the best match-score is used for contig classification. Contigs 
are classified as ‘Prophage-like’, ‘Sloping’, or 'NoPattern' during 
pattern-matching. 

#### **Patterns**

##### Sloping:
There are four sloping pattern variations in the sloping pattern class. The 
sloping patterns are representative of large DNA transfers that take place 
during generalized, lateral and gene transfer agent (GTA) transduction due to 
the decreasing frequency of DNA packaging moving away from the packaging 
initiation sites. During pattern-matching, the slope values of the sloping 
patterns are decreased until a minimum slope of 0.001 (change of 10x read 
coverage over 10,000 bp) is reached. The minimum slope value can be changed 
with the `minSlope` parameter. Generalized, lateral and GTA transduction events 
can span tens to hundreds of kilobasepairs of DNA and a single contig typically 
does not capture an entire event. Depending on which part of the transducing 
event is captured by the contig, the sloping can be very severe or almost 0. 
Patterns 1 and 2 represent contigs that capture a Sloping transducing event 
somewhere in the middle of the DNA transfer. Patterns 2 and 4 represent contigs 
that capture the jump of read coverage associated with packaging initiation 
site of a Sloping transducing event. Patterns 2 and 4 are translated across the 
contig in addition to having the slopes changed while only the slopes are 
changed on patterns 1 and 2.

```{r echo=FALSE, fig.width=4,fig.height=4}
dataframe <- cbind.data.frame(c(1:100), c(seq(1, 100, 1)))
colnames(dataframe) <- c("mockpos", "mockcov")
plot1 <- ggplot(dataframe, aes(x = mockpos, y = mockcov)) +
  geom_line(linewidth = 1) +
  labs(x = NULL, y = NULL, title = "Pattern 1") +
  theme_classic() +
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(size = 10),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2)
  )

dataframe <- cbind.data.frame(c(1:100), c(seq(100, 1, -1)))
colnames(dataframe) <- c("mockpos", "mockcov")
plot2 <- ggplot(dataframe, aes(x = mockpos, y = mockcov)) +
  geom_line(linewidth = 1) +
  labs(x = NULL, y = NULL, title = "Pattern 2") +
  theme_classic() +
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(size = 10),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2)
  )

dataframe <- cbind.data.frame(c(1:100), c(rep(0, 20), seq(80, 1, -1)))
colnames(dataframe) <- c("mockpos", "mockcov")
plot3 <- ggplot(dataframe, aes(x = mockpos, y = mockcov)) +
  geom_line(linewidth = 1) +
  labs(x = NULL, y = NULL, title = "Pattern 3") +
  theme_classic() +
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(size = 10),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2)
  )

dataframe <- cbind.data.frame(c(1:100), c(seq(1, 80, 1), rep(0, 20)))
colnames(dataframe) <- c("mockpos", "mockcov")
plot4 <- ggplot(dataframe, aes(x = mockpos, y = mockcov)) +
  geom_line(linewidth = 1) +
  labs(x = NULL, y = NULL, title = "Pattern 4") +
  theme_classic() +
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(size = 10),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2)
  )

plot1 + plot2 + plot3 + plot4
```

##### Prophage-like:
There are three block patterns in the Prophage-like pattern class. The block 
patterns are representative of integrated genetic elements that can be excised 
from the host chromosome and mobilized. The blocks of read coverage that define 
this classification are formed when the reads of mobilized genetic elements 
packaged by VLPs map back to their respective integration sites in the host 
bacterium's chromosome. Prophage, phage-inducible chromosomal islands (PICIs), 
and transposons are all examples of genetic elements that fall into the 
Prophage-like class. During pattern-matching, both the heights and widths of 
block patterns are altered and all pattern variations are translated across the 
contig. The block pattern widths never get smaller than 10,000 bp by default, 
however this can be changed with the `minBlockSize` parameter. Pattern 1 
represents a Prophage-like element that is entirely on the contig while 
patterns 2 and 3 represent Prophage-like elements that trail off the right or 
left side of the contig, respectively. While a Prophage-like classification is 
not an example of transduction by itself, there may be transduction associated 
with Prophage-like classifications. The improper excision of Prophage-like 
elements may lead to the specialized transduction of genes neighboring the 
element's integration site. TrIdent's `specializedTransductionID()` function 
(explained in detail below) will attempt to detect specialized transduction 
events associated with Prophage-like classifications. 

```{r echo=FALSE, fig.width=4,fig.height=2}
dataframe <- cbind.data.frame(c(1:100), c(rep(0, 20), rep(100, 60), rep(0, 20)))
colnames(dataframe) <- c("mockpos", "mockcov")
plot1 <- ggplot(dataframe, aes(x = mockpos, y = mockcov)) +
  geom_line(linewidth = 1) +
  labs(x = NULL, y = NULL, title = "Pattern 2") +
  theme_classic() +
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(size = 10),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2)
  )

dataframe <- cbind.data.frame(c(1:100), c(rep(0, 60), rep(100, 40)))
colnames(dataframe) <- c("mockpos", "mockcov")
plot2 <- ggplot(dataframe, aes(x = mockpos, y = mockcov)) +
  geom_line(linewidth = 1) +
  labs(x = NULL, y = NULL, title = "Pattern 3") +
  theme_classic() +
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(size = 10),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2)
  )

dataframe <- cbind.data.frame(c(1:100), c(rep(100, 40), rep(0, 60)))
colnames(dataframe) <- c("mockpos", "mockcov")
plot3 <- ggplot(dataframe, aes(x = mockpos, y = mockcov)) +
  geom_line(linewidth = 1) +
  labs(x = NULL, y = NULL, title = "Pattern 1") +
  theme_classic() +
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(size = 10),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2)
  )

plot3 + plot1 + plot2
```


##### noPattern:
Since the best pattern-match for each contig is determined by comparing 
match-scores amongst all pattern-variations from all pattern classes, we needed 
a ‘negative control’ pattern to compare against. The 'NoPattern' 'pattern' 
serves as a negative control by matching to contigs with no read coverage 
patterns. We made two NoPattern patterns which consist of a horizontal line the 
same length as the contig being assessed at either the average or median read 
coverage for a contig. This pattern is not re-scaled or translated in any way. 
Note that read coverage patterns are heavily dependent on the depth of read 
coverage achieved during sequencing and therefore very rare transduction events 
may not achieve sufficient read coverage for detection with read coverage 
pattern-matching. Rather than label contigs with no read coverage pattern as 
having ‘no transduction’, we instead label them as having 'no pattern'.

```{r echo=FALSE, fig.width=4,fig.height=4}
dataframe <- cbind.data.frame(c(1:100), rep(10, 100))
colnames(dataframe) <- c("mockpos", "mockcov")
plot1 <- ggplot(dataframe, aes(x = mockpos, y = mockcov)) +
  geom_line(linewidth = 1) +
  ylim(0, 100) +
  labs(x = NULL, y = NULL, title = "Pattern 1") +
  theme_classic() +
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(size = 10),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2)
  )
plot1
```


### Highly active/abundant and heterogenously integrated/present Prophage-like elements

Prophage-like elements that are actively replicating or are highly abundant 
will typically generate more sequencing reads than the rest of their host 
bacterium's chromosome. This may create a region of elevated read coverage at 
the element's insertion site in the whole-community fraction read coverage. 
Conversely, if a Prophage-like element is integrated into only a portion of the 
host bacterial population, there may be a dip or depression in read coverage at 
the integration site in the whole-community read coverage. In order to 
determine if the whole-community read coverage is elevated or depressed at the 
site of a Prophage-like element, one must know the Prophage-like element's 
genomic location. While some tools rely on annotation information to identify 
Prophage-like elements in whole-community metagenomes, `TrIdentClassifier()` 
uses the VLP-fraction read coverage patterns. The locations of Prophage-like 
pattern-matches are used to calculate the Prophage-like:non-Prophage-like 
whole-community read coverage ratio. Prophage-like patterns with 
whole-community read coverage ratios greater than 1.15 are labeled as 'elevated' 
while ratios less then 0.75 are labeled as 'depressed'. 

### NoPattern classifications with high VLP-fraction:whole-community read coverage ratios

If a contig receives a noPattern classification, it proceeds to an additional 
classification step which either leaves the classification as is or 
re-classifies the contig as having a high VLP-fraction:whole-community read 
coverage ratio (‘HighCovNoPattern’). In other words, contigs with the 
HighCovNoPattern classification have even read coverage across the contig 
(i.e. no notable read coverage pattern) and the VLP-fraction read coverage is 
notably higher than the whole-community read coverage. This re-classification 
is necessary because certain biologically significant events that may be of 
interest to the user can have these characteristics. First, certain 
transduction events, specifically those associated with the Sloping 
classification, may be so long that they extend across multiple contigs. The 
‘tails’ of the sloping patterns created by these events may have little to no 
actual sloping (no pattern) but relatively high read coverage (Fig S2 Kleiner 
et al., 2020). Second, there is evidence that membrane vesicle (MV) mediated 
transduction (also known as vesiduction or protected extracellular 
transformation) produces relatively even read coverage patterns when purified 
MV sequencing reads are mapped back to their bacterial chromosome of origin 
(Faddetta et al., 2022). Lastly, phage genomes that have assembled into contigs 
in the whole-community fraction may generate high levels of even read coverage 
as the phage reads in the VLP-fraction map back to their own genome sequences. 
Contigs with median VLP-fraction:Whole-community read coverage ratios greater 
than 2, in other words contigs where the median VLP-fraction read coverage 
value is 2x the whole-community median read coverage value, are re-classified 
as HighCovNoPattern.

**Reference:** Faddetta, T., Vassallo, A., Del Duca, S. et al. Unravelling the 
DNA sequences carried by Streptomyces coelicolor membrane vesicles. Sci Rep 12, 
16651 (2022). <https://doi.org/10.1038/s41598-022-21002-z>

## Usage

Default arguments:
```{r}
TrIdentOutput <- TrIdentClassifier(
  VLPpileup = VLPFractionSamplePileup,
  WCpileup = WholeCommunitySamplePileup
)
```

## Arguments/Parameters
```{r eval=FALSE}
TrIdentClassifier(VLPpileup, WCpileup,
  windowSize = 1000, minBlockSize = 10000,
  maxBlockSize = Inf, minContigLength = 30000, minSlope = 0.001,
  suggFiltThresh = FALSE, verbose = TRUE, SaveFilesTo
)
```

* `VLPpileup`: VLP-fraction pileup file. 
* `WCpileup`: Whole-community pileup file.
* `windowSize`: The number of basepairs to average read coverage values over. 
Options are 100, 200, 500, 1000 ONLY.  Default is 1000.
* `minBlockSize`: The minimum size (in bp) of the Prophage-like block pattern. 
Default is 10000. Must be greater than 1000.
* `maxBlockSize`: The maximum size (in bp) of the Prophage-like block pattern. 
Default is NA (no maximum).
* `minContigLength`: The minimum contig size (in bp) to perform 
pattern-matching on. Must be at least 25000. Default is 30000.
* `minSlope`: The minimum slope value to test for sloping patterns. Default is 
0.001 (i.e minimum change of 10x read coverage over 100,000 bp).
* `suggFiltThresh`: TRUE or FALSE, Suggest a filtering threshold for TrIdent 
classifications based on the normalized pattern-match scores. Default is FALSE.
* `verbose`: TRUE or FALSE. Print progress messages to console. Default is TRUE.
* `SaveFilesTo`: Optional, Provide a path to the directory you wish to save 
output to. A folder will be made within the provided directory to store 
results.

## Output

The output of `TrIdentClassifier()` is a list containing five objects:

1.  SummaryTable: A table containing the classification information for all 
contigs that were not filtered out.
2.  CleanedSummaryTable: A cleaned summary table containing the classification 
information for all contigs classified 
as Prophage-like, Sloping and HighCovNoPattern.
3.  PatternMatchInfo: A list of pattern-match information that is used by other 
functions in TrIdent.
4.  FilteredOutContigTable: A table containing contigs that were filtered out 
and the reason why (low read coverage or too short).
5.  windowSize: The `windowSize` used.
6.  ResultHistogram: a histogram displaying the overall abundance and 
quality of pattern-matches in addition to the composition of classifications. 
The displayed pattern-match scores are normalized by dividing each score by its 
associated contig length. The scores are normalized to visualize the overall 
quality of pattern-matching for the entire dataset. Remember, smaller 
pattern-match scores correspond to better pattern-matches. 

Save the desired list-item to a new variable using its associated name.

Summary table:
```{r}
TrIdentSummaryTable <- TrIdentOutput$SummaryTable
```

```{r, echo=FALSE}
kable(TrIdentSummaryTable) %>%
  kable_styling(latex_options = "HOLD_position")
```

# plotTrIdentResults()

`plotTrIdentResults()` allows users to visualize both the whole-community and 
VLP-fraction read coverage and the pattern-match associated with each contig 
classified as Prophage-like, Sloping and HighCovNoPattern.  

## Function components

### Re-building pattern-matches
The `TrIdentClassifier()` output contains information needed to re-build each 
pattern-match used for contig classification. To re-build a complete 
pattern-match for visualization, `plotTrIdentResults()` uses the 
pattern-match's minimum and maximum values and the start and stop positions.

### Plotting read coverage and associated pattern-matches
The whole-community and VLP-fraction read coverage are plotted for each contig 
classified as Prophage-like, Sloping or HighCovNoPattern. The pattern-match 
associated with the classification is overlaid on the VLP-fraction read 
coverage. The whole-community read coverage is displayed for visualization of 
elevated or depressed Prophage-like region read coverage. 

## Usage

Default arguments:
```{r}
TrIdentPlots <- plotTrIdentResults(
  VLPpileup = VLPFractionSamplePileup,
  WCpileup = WholeCommunitySamplePileup,
  TrIdentResults = TrIdentOutput
)
```

## Arguments/Parameters
```{r eval=FALSE}
plotTrIdentResults(
  VLPpileup, WCpileup, TrIdentResults, matchScoreFilter,
  saveFilesTo
)
```

* `VLPpileup`: VLP-fraction pileup file.
* `WCpileup`: Whole-community pileup file.
* `TrIdentResults`: The output from `TrIdentClassifier()`
* `matchScoreFilter`: Optional, Filter plots using the normalized pattern 
match-scores. A suggested filtering threshold is provided by 
`TrIdentClassifier()` if `suggFiltThresh=TRUE`.
* `saveFilesTo`: Optional, Provide a path to the directory you wish to save 
output to. A folder will be made within the provided directory to store 
results.

## Output

The output of `plotTrIdentResults()` is a list containing ggplot objects. The 
list contains all read coverage plots for contigs classified as Sloping, 
Prophage-like, or HighCovNoPattern and their respective pattern-matches.

**By default, the plots are displayed with raw read coverage values. We 
recommend that users also view plots using `logScale=TRUE` as some specialized 
transduction patterns occur at such low frequencies they can only be visualized
using log scaled read coverage values.** 


View select plots:
```{r}
TrIdentPlots$NODE_62
TrIdentPlots$NODE_368
TrIdentPlots$NODE_352
```

# specializedTransductionID()

Specialized transduction occurs when an integrated Prophage-like element both 
excises and packages a small portion of bacterial DNA directly outside the 
borders of its insertion site. `specializedTransductionID()` searches contigs 
classified as Prophage-like for dense read coverage outside the borders of the 
Prophage-like patterns that might represent specialized transduction events. 
Because specialized transduction tends to be fairly short (a couple kbps) 
compared to generalized, lateral or GTA transduction (tens to hundreds of 
kbps), averaging read coverage over a 1,000 bp distance (i.e using a 
`windowSize=1000`) can 'blur' specialized transduction patterns. This is why 
specialized transduction is not identified in `TrIdentClassifier()`. Instead, 
we use the locations of Prophage-like elements identified with 
`TrIdentClassifier()` to guide our search for specialized transduction in 
`specializedTransductionID()`.

Note that `specializedTransductionID()` results should not be taken at face 
value and some manual curation is needed! We recommend careful examination of 
the output plots to determine if you agree with the 
`specializedTransductionID()` results. Setting `logScale=TRUE` will plot the 
read coverage in log10 scale and may help with visualization of low-frequency 
specialized transduction events. Be aware that in some cases, log scaled read 
coverage can obfuscate the Prophage-like element pattern itself.

## Function components

### Zoom-in on Prophage-like elements
Since, specialized transduction read coverage patterns are relatively small, 
`specializedTransductionID()` 'zooms-in' on Prophage-like elements to aid with 
specialized transduction visualization. To 'zoom-in' on Prophage-like elements, 
`specializedTransductionID()` uses the pattern-match information generated in 
`TrIdentClassifier()` to subset the contig 50,000 bp outside the pattern-match 
borders.

### Identify borders of Prophage-like elements
Since the `windowSize` used for pattern-matching in `TrIdentClassifier()` is 
*likely* larger than 100 bp, the Prophage-like pattern-match border locations 
will not directly translate to the border locations at the 100 bp `windowSize`. 
To re-identify the borders of Prophage-like elements, we use a rolling standard 
deviation (SD) to detect large changes in read coverage associated with 
Prophage-like element borders. First, the contig is subset surrounding the left 
and right pattern-match borders. For contigs that have a Prophage-like 
pattern-match that trails off the side of a contig, then only the border that 
falls on the contig is subset. Next, for each subset, the rolling SD is 
calculated between each read coverage value and the contig position associated 
with the largest SD is chosen as the border location. The border locations are 
marked on the output plots with black vertical lines.

### Search for specialized transduction outside Prophage-like borders
Once the Prophage-like element borders are identified, 
`specializedTransductionID()` starts from the border locations and searches 
outwards for dense read coverage that meet the 'requirements' for specialized 
transduction as defined by two parameters in `specializedTransductionID()`:

* `noReadCov`
* `specTransLength`

`specializedTransductionID()` first makes sure that any coverage it detects 
outside the borders is not disrupted by a defined region of no read coverage 
(`noReadCov`). The default value for `noReadCov` is 500 bp. Secondly, 
`specializedTransductionID` ensures that any read coverage it detects outside 
of the Prophage-like borders meets a minimum length requirement 
(`specTransLength`). The default value for `specTransLength` is 2000 bp. So by 
default, `specializedTransductionID()` will search for coverage directly 
outside of the Prophage-like borders that is at least 2000 bp long and is not 
interrupted at any point by more than 500 bp of no read coverage. If these 
requirements are met, `specializedTransductionID()` will label the contig as 
having specialized transduction. The 'end' of the specialized transduction 
region is marked on the output plots with red vertical lines. Additionally, 
read coverage plots for contigs with specialized transduction are colored green 
instead of blue. 

## Usage

Default arguments:
```{r}
SpecTransduc <- specializedTransductionID(
  VLPpileup = VLPFractionSamplePileup,
  TrIdentResults = TrIdentOutput
)
```

## Arguments/Parameters
```{r eval=FALSE}
specializedTransductionID(VLPpileup, TrIdentResults,
  noReadCov = 500, specTransLength = 2000,
  logScale = FALSE, verbose = TRUE, matchScoreFilter,
  SaveFilesTo, specificContig
)
```

* `VLPpileup`: VLP-fraction pileup file.
* `TrIdentResults`: The output from `TrIdentClassifier()`.
* `noReadCov`:  Number of basepairs of zero read coverage encountered before 
specialized transduction searching stops. Default is 500. Must be at least 100.
* `specTransLength`: Number of basepairs of non-zero read coverage needed for 
specialized transduction to be considered. Default is 2000. Must be at least 100.
* `logScale`: TRUE or FALSE, display VLP-fraction read coverage in log10 scale. 
Default is FALSE.
* `matchScoreFilter`: Optional, Filter plots using the normalized pattern 
match-scores. A suggested filtering threshold is provided by 
`TrIdentClassifier()` if `suggFiltThresh=TRUE`. 
* `verbose`: TRUE or FALSE. Print progress messages to console. Default is TRUE.
* `SaveFilesTo`: Optional, Provide a path to the directory you wish to save 
output to. A folder will be made within the provided directory to store results.
* `specificContig`: Optional, Search a specific contig classified as 
Prophage-like ("NODE_1").

## Output

The output of `specializedTransductionID()` is a list containing two objects:

1. summaryTable: A table containing the `specializedTransductionID()` results 
for all contigs classified as Prophage-like in `TrIdentClassifier()`.
2. Plots: A list of ggplot objects for all contigs classified as Prophage-like. 
The plots are 'zoomed-in' on the Prophage-like pattern to aid with specialized 
transduction visualization. The Prophage-like element borders identified by 
`specializedTransductionID()` are marked on each plot with black vertical 
lines. If `specializedTransductionID()` identifies potential specialized 
transduction, the plot will be colored green instead of blue. The end of 
specialized transduction will be marked with a red vertical line.

Save the desired list-item to a new variable using its associated name.

Summary table:
```{r}
SpecializedTransducSummaryTable <- SpecTransduc$summaryTable
```

```{r, echo=FALSE}
kable(head(SpecializedTransducSummaryTable), format = "html") %>%
  kable_styling(latex_options = "HOLD_position")
```

View select plots:
```{r}
SpecTransduc$Plots$NODE_62
SpecTransduc$Plots$NODE_135
```

# Generate final TrIdent summary table

If you'd like to combine the summary tables produced by `TrIdentClassifier()` 
and `specializedTransductionID()`, try the following:
```{r}
FinalTrIdentSummaryTable <- merge(TrIdentSummaryTable,
  SpecializedTransducSummaryTable,
  by = "contigName", all.x = TRUE
)
```

```{r echo=FALSE}
kable(FinalTrIdentSummaryTable) %>%
  kable_styling(latex_options = "HOLD_position")
```

# Supplemental information

## Acknowledgments
TrIdent is the result of the combined effort and brain power of many 
individuals. Specifically, we would like to thank Dr. Manuel Kleiner, 
Dr. Ben Callahan, Dr. Breck Duerkop and Dr. Craig Gin for their individual 
expertise and overall support!

## Funding
The development of TrIdent was supported by a seed grant from the North 
Carolina State University Data Science Academy and by the National Institutes 
of Health under Award Numbers R35GM138362 and R01Al171046.

## Session Information
```{r}
sessionInfo()
```
